---
title: "Sampling"
---

Follow along with the demo to learn how to sample data from a population using R and the Tidyverse. You should get started by loading the tidyverse in a new R block.

### Task 1: Load the tidyverse
```{r}
library(tidyverse)
# i forgot that i could comment out whatever i want 
```


### Task 2: Load 1,000 rows of data with the ages of people from 18-80
```{r}
set.seed(1114)
# can be any number, has to be a number (42, 16, 123, etc.)

population <- tibble(
  age = sample(18:80, 1000, replace = TRUE)
)
# "replace" means to overwrite what we previously have, "sample" is the range of numbers and the other number is how many of them (right...? i kinda missed it)
```


### Task 3: Sample 100 rows of data from the dataset
```{r}
sample100 <- population |>
  sample_n(100)
# now we are sampling 100 rows from our population dataframe
# oh okay so every time you run, it'll take a different sample set
```


### Task 4: Calculate the mean age and standard deviation of the sample
```{r}
sample100 |>
  summarize(
    mean_age = mean(age),
    sd_age = sd(age)
  )
# taking mean and standard deviation
# so we all put different numbers, but we're working in the same range so they sould be similar (which, they were)
```


### Task 5: Calculate the mean age and standard deviation of the population
```{r}
population |>
  summarize(
    mean_age = mean(age),
    sd_age = sd(age)
  )
# same thing as task 4 but now with population
```


### Task 6: Compare the sample and population statistics

Compare the means and standard deviations. Are they similar? Go back and re-run the sampling and statistics calculations with a smaller sample size. How do the results change?

Yes, the means and standard deviations of sample100 and population are very similar, and are only a few numbers off very frequently. Sometimes it's slightly off, I got 45 for the mean at one point, but the rest of them stayed somewhere between 48-51 for their mean. The standard deviation also stays between 17-19. So, the results are often close.


When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
