---
title: "UMD Fees"
---

Follow along with the demo to learn how to calculate percentage change using R and the Tidyverse. You should get started by loading the tidyverse in a new R block.


### Task 1: Load the tidyverse


```{r}
library(tidyverse)
```




### Task 2: Load the UMD Fees dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/umd_fees.csv

```{r}
umd_fees <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/umd_fees.csv")
```

### Task 3: Calculate the percentage change in each category between Fall 2026 and Fall 2021 using `mutate()`

```{r}
umd_fees_pct <- umd_fees |> mutate(pct_change = (`Fall 2026`- `Fall 2021`)/`Fall 2021` * 100)
```



### Task 4: Describe the results

Write a few sentences about what you found in those percentages. What's the story here? What questions do you have?

The largest percentage changes come from Student Counseling Center and Student Sustainability. However, technically the Student Counseling Center percentage change isn't measurable, because it starts at zero. It most likely didn't even exist yet, and now there is a fee because it exists. My theory is that there were probably a few events like the pandemic and the stadium incident that spurred the school to implement this fee.

Other than that, Student Sustainability and Shuttle Bus fees are the highest percentage changes. Student Sustainability is not as big a fee as the other fees, and its high percentage change would most likely not be notable to most people because it is not a three-digit number like most of the others. Though, I am not totally sure what "Student Sustainability" wholly includes, so I have questions about what it entails and then I would see how much people really get use out of it, if they are even aware of it in the first place.

For the Shuttle Bus, I have questions about who uses the bus more, whether it's people living on campus or off campus, grade levels, etc. I also have questions about whether this money supports the bus fuel or the bus upkeep, for example, if a bus got damaged, are these fallback funds for them?


When you are done, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
